---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
# the search endpoints are throttled to two requests per minute, which httr2
# reports with a progress bar that has no business in a rendered README
options(cli.progress_show_after = Inf)
```

# mediacloud2 <img src="man/figures/logo.svg" align="right" height="139" alt="" />

<!-- badges: start -->
[![R-CMD-check](https://github.com/JBGruber/mediacloud2/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/JBGruber/mediacloud2/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/JBGruber/mediacloud2/graph/badge.svg)](https://app.codecov.io/gh/JBGruber/mediacloud2)
<!-- badges: end -->

The goal of mediacloud2 is to make the [Media Cloud](https://www.mediacloud.org) news archive usable from R.
Media Cloud collects online news from tens of thousands of outlets worldwide and lets you search that archive by keyword, outlet, country and date.
The package wraps version 4 of the Media Cloud API, which replaced the retired v2 API that older R clients were written against.

Everything the package returns is a tibble, so results go straight into your usual data wrangling and plotting code.

## Installation

You can install the development version of mediacloud2 like so:

``` r
pak::pak("https://github.com/JBGruber/mediacloud2")
```

# Basic Usage

```{r}
library(mediacloud2)
```

## Authentication

To use the MediaCloud API, you first need to obtain a token.
Head to <https://search.mediacloud.org/account> and create an account, if you do not have one already, then create a token.
Or you skip that step and run `mc_auth()` directly:

```
mc_auth()
```

This will open the URL you need and ask you to enter the token.
This way, the token doesn't show up in your R history or anywhere else.
After you have done this once, the package will never ask you again to authenticate, unless you change computers or delete the cache directory.
You can learn more about the function with `?mc_auth()`.

## Finding Sources

Media Cloud sorts the outlets it archives into *sources* -- one per domain -- and *collections*, which are named groups of sources, mostly countries and regions.
Every search runs against at least one of the two, so this is where you normally start.

To find an individual source, search for its name:

```{r}
spiegel <- mc_sources(name = "spiegel")
spiegel[, c("id", "name", "stories_per_week", "primary_language")]
```

Searching for a country instead gives you the collections it is covered by:

```{r}
mc_find_collections("Germany")
```

`"Germany - National"` is the one you usually want.
Its id, `34412409`, is what the search functions below take, but you can also ask for the sources it contains:

```{r}
sources_de <- mc_collection_sources("Germany - National")
sources_de[, c("id", "name", "stories_per_week", "last_story")]
```

## Getting News Items

`mc_story_list()` returns the stories matching a query.
The query goes to Elasticsearch, so the usual `AND`/`OR`/`NOT` operators, quoted phrases and wildcards all work.
Search terms are always paired with a date range and with at least one collection or source:

```{r}
stories <- mc_story_list(
  "bluesky",
  start_date = "2026-01-01",
  end_date = "2026-09-01",
  collection_id = 34412409
)
stories
```

Paging is handled for you: the function keeps asking for the next page until the archive is exhausted, or until `max_results` is reached.
Media Cloud allows ordinary accounts two search requests per minute, and the package waits rather than running into that limit, so a broad query takes a while.

If you only want the shape of the coverage rather than the stories themselves, `mc_count_over_time()` answers in a single request:

```{r bluesky-over-time, fig.height = 3.5, fig.width = 8, dpi = 150, fig.alt = "Bar chart of the number of stories per day in German national news that mention Bluesky, January to September 2026."}
counts <- mc_count_over_time(
  "bluesky",
  start_date = "2026-01-01",
  end_date = "2026-09-01",
  collection_id = 34412409
)
counts

library(ggplot2)
ggplot(counts, aes(x = date, y = count)) +
  geom_col(fill = "#2a5db0", width = 1) +
  labs(
    title = "German national news mentioning Bluesky",
    x = NULL,
    y = "Stories per day"
  ) +
  theme_minimal()
```

## Obtaining Tokens

MediaCloud does not give you the raw text of a news item, but it lets you get the individual tokens or features of a story.
That means you get all the words that were originally in a story, but not the order of the words.

```{r}
words <- mc_words(
  "bluesky",
  start_date = "2026-01-01",
  end_date = "2026-09-01",
  collection_id = 34412409
)
words
```

The counts come from a sample of the matching stories, not from all of them, which is why `sample_size` is reported alongside them.

You can use these words, for example, to build a word cloud with `ggwordcloud`.

```{r bluesky-wordcloud, fig.height = 5, fig.width = 8, dpi = 150, fig.alt = "Word cloud of the terms most common in German national news stories mentioning Bluesky, with linke, social, gruene, berlin and spd the largest."}
library(ggwordcloud)
ggplot(words, aes(label = term, size = term_count, colour = term_count)) +
  geom_text_wordcloud(seed = 1, grid_margin = 4) +
  scale_size_area(max_size = 14) +
  scale_colour_gradient(low = "#7ba7e3", high = "#12305e") +
  theme_minimal()
```

# Alternatives

There have been other R packages in the past ([`mediacloudr`](https://github.com/jandix/mediacloudr) by jandix, [`joon-e/mediacloud`](https://github.com/joon-e/mediacloud)), however, they target the retired v2 API and are unusable against the current v4.
So the only working alternative, as far as I can tell, is the [Python package by MediaCloud themselves](https://github.com/mediacloud/api-client).

# How to cite

**This R Package**:

> Gruber, J. B. (2026). *mediacloud2: MediaCloud API (v4) Wrapper*. R package version 0.0.0.9000. <https://github.com/JBGruber/mediacloud2>

**The API**:

> Bermejo, F., Bhargava, R., Budne, P., Gulley, P., Leon, E., McGrady, R., … Zuckerman, E. (2026). Media Cloud 2.0: An Updated Open Web News Archive. Proceedings of the International AAAI Conference on Web and Social Media, 20(1), 2735–2746. https://doi.org/10.1609/icwsm.v20i1.42778
